\documentclass{article}
\pagestyle{empty}
\usepackage{amsmath}
\usepackage{amsthm}

\begin{document}

%% Files needed for submittal:
% bp2.m
% bp2val.m
% Anderson_hw6_p1.m
% prices.mat
% input.mat
% msa_voronoi_quant.m
% Anderson_hw6_p2.m
% hw6_train_gen.m

\flushright
Mathew Anderson \\
Sys Eng 5212 Intro to Neural Networks and Applications, Spring 2015 \\ HW 6

\flushleft
1. For this homework, I used weather data for five cities in Iowa
to train a neural network using backpropagation.
The date range of my weather data spans from April 1985 to December 1999.
My output data is historical monthly corn prices as measured in
US\$ per Bushel according to http://www.indexmundi.com/commodities/?commodity=corn\&months=360.
The output quantization is the delta of the monthly price normalized
to $[-0.8, 0.8]$.
The five cities selected as ``weather stations'' are:
\begin{itemize}
\item Algona
\item Ames
\item Atlantic
\item Audubon
\item Boone
\end{itemize}
These cities were chosen primarily because of data availability and
their approximate geographic locations.
Cities in the state of Iowa were selected because Iowa produces more corn
than any other state in the United States and the United States is the leading
corn producing country in the world.

For each day in each month, the daily summary of weather data, as provided
by the Weather Underground (www.wunderground.com), gets specific features
(average barometric pressure, average daily temperature, total precipication)
classified according to the following
``buckets'':
\begin{itemize}
\item Pressure (in inHg): $[29.4, 29.5, 29.6, 29.7, 29.8, 29.9, 30, 30.1,
30.2, 30.3, 30.4, 30.5]$
\item Temperature (in $^{\circ}F$): $[0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110]$
\item Precipitation (in inches): $[0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,
2.25, 2.5, 2.75, 3]$
\end{itemize}
Note that any values in excess of the highest bucket or less than the value
of the lowest bucket will result in a quantization of 12 or 0.
No scientific rationale for these buckets exist, they were chosen to match
http://omahanebraskawx.com/wxbarodetail.php.\\

\vspace{6pt}
The network itself consists of 5 hidden neurons with one output neuron.
Each neuron will have 420 inputs, representing 28 samples of 3 data features
for each of the 5 cities.
28 samples are used to represent a month for simplicity.

\vspace{6pt}
Because the weather data and monthly pricing data spans 177 months,
the first 123 input samples will serve as training data and the final
54 samples will serve as testing data.
The input data will be shuffled before the backpropagation algorithm
is allowed to execute.
\vspace{6pt}

See Anderson\_hw6\_p1.m for the final implementation of my network.
An MSE plot is produced upon running the script.

\vspace{6pt}
The backpropagation algorithm uses Nguyen-Widrow weight initialization and
fits reasonably well ``out of the box'' using a mommentum paramter of
0.01.
I noticed that overfitting appears to occur rapidly (before 100 iterations),
so I tuned the cutoff for the growth of the MSE to 0.0001.
The number of hidden neurons was determined experimentally from the following
set $[3, 5, 10]$.
Five most consistently produced MSE values between 0.02 and 0.03.

\vspace{6pt}
Further refinements would have to originate in the data.
Unfortunately, Weather Underground doesn't have good pressure or precipitation
readings before the mid 1990s for most cities in Iowa.
Other websites contain more complete weather data but without the historical
depth of Weather Underground.
As an added bonus, most non-Weather Underground websites are free.
It's possible that the ``holes'' in the barometric pressure and
precipitation data are the cause of my relatively high MSE.
I consider that a possibility because, for example, without adequate precipitation
data it will not be possible to detect floods/droughts.

\vspace{6pt}
Finally, there are many varieties of corn. While the United States is the leading corn
producing nation and Iowa produces more corn than any other state, both Minnesota
and Illinois produce a large amount of corn per year as well and I expect that weather
conditions in the neighboring states also affect the price of corn per month.
The weather is not an independent thing and the ``Corn Belt'' in the United States
consists primarily of Iowa, Minnesota, South Dakota, and Illinois.
If it is possible to maintain the number of cities classified as ``weather stations'' while
expanding the geographic coverage, it may certainly improve the relationship between weather
observations and change in corn as measured in US\$ per bushel.
\vspace{6pt}

2. See Anderson\_hw6\_p2.m, hw6\_train\_gen.m, and msa\_voronoi\_quant.m
for my solution.
The main script (Anderson\_hw6\_p2.m) will produce a figure that proves
the training data matches the problem statement and one figure of
classification output for various centroid counts.
Higher centroid counts improve classification accuracy,
higher training iterations.

\end{document}
